GPU Acceleration of the Material Point Method

### 2019-02-10 pres/pres.tex
motiv., begin math

\documentclass{beamer}
\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{mathtools}
\usepackage{subdepth}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{svg}
\usepackage{xcolor,colortbl}

@@

@@

\begin{document}
\begin{frame}
  \begin{enumerate}
    \item \cite{MPM:COURSE}
  \end{enumerate}
\end{frame}
\selectlanguage{english}
\printbibliography
\end{document}

>>>>

\documentclass{beamer}
\usetheme{Malmoe}
\usecolortheme{seahorse}

\usepackage{textpos}
\usepackage{csquotes}
\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{mathtools}
\usepackage{subdepth}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{svg}
\usepackage{xcolor,colortbl}
\usepackage{media9}

\let\oldcite=\cite
\renewcommand{\cite}[1]{\textcolor[rgb]{.55,.55,.89}{\oldcite{#1}}}

\setbeamerfont{frametitle}{size=\small}

@@

\addbibresource{pres.bib}

@@

%Information to be included in the title page:

\title{GPU Acceleration of the Material Point Method}

\author % (optional, for multiple authors)
{Fabian Meyer}

\institute % (optional)
{University of Koblenz}

\date{Kolloquium Computergrafik, 14 February 2019}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\section{Motivation on a GPU MPM Approach}

\subsection{A Brief MPM Overview: Do You Want to Build a Snowman?}

\begin{frame}

\frametitle{A Brief MPM Overview: Do You Want to Build a Snowman?}
A short historical summary of MPM:
\vfill
\begin{minipage}{0.43\paperwidth}
  \begin{itemize}
    \item Belongs to family of particle-in-cell(PIC) techniques \cite{evans1957particle}.
    \item Initial application to solids \cite{sulsky1995application} $\rightarrow$ MPM
    \item From research to production in \textit{Disney's} animation film \textit{Frozen} \cite{MPM:SNOW}.
    \item Avalanche research \cite{MPM:AVALANCHE}
  \end{itemize}
\end{minipage}
\begin{minipage}{0.35\paperwidth}
\includemedia[
  width=144,
  height=81,
  activate=pageopen,
  addresource=media/snowh264.mp4,
  flashvars={source=media/snowh264.mp4&autoPlay=true&loop=true&repeat=always}
  ]{\includegraphics{media/snow.jpg}}{VPlayer.swf}
  \tiny{Video result of my bachelor thesis on the simulation of snow \cite{Meyer2015}.}
\end{minipage}
\end{frame}

\begin{frame}
PIC ideas:
\vfill
\begin{minipage}{0.45\paperwidth}
\begin{itemize}
    \item Combine Lagrangian particles \& Eulerian grid    \item Particles store all information
\end{itemize}
\vfill
Typical PIC/MPM roundtrip:
\vfill
\begin{enumerate}
  \item Particle-to-grid(P2G) transfer to an unmoving grid
    \item Solve discretized governing equations on grid
    \item Grid-to-particle(G2P) transfer back to particles \& move them
    \item[$\Rightarrow$] \textbf{meshfree, non-empirical}
\end{enumerate}
\end{minipage}
\begin{minipage}{0.35\paperwidth}
	\includesvg[width=0.35\paperwidth]{media/pic.svg}
	\tiny{Transfers: Interpolation functions are defined over grid nodes.}
\end{minipage}
\end{frame}
\subsection{GPGPU for performance enthusiasts}
\begin{frame}
  \frametitle{GPGPU for performance enthusiasts}
  Why would('nt) you?
  \vfill
  \begin{columns}[T]
    \begin{column}{0.45\textwidth}
      \textbf{Drawbacks:}
    \begin{itemize}
      \item Interactivity much easier on CPU, but slow PCI-Bus communication
      \item Code is mostly written against GPU architecture
      \item A lot of strain on the programmer
    \end{itemize}
    \end{column}
    \begin{column}{0.45\textwidth}
      \textbf{Benefits:}
    \begin{itemize}
      \item Data is already on the GPU for rendering
      \item \textbf{Higher parallelization acceleration}
    \end{itemize}
    \end{column}
  \end{columns}
\end{frame}
\section{A Gentle Introduction to the MPM}

\subsection{Governing Equations: Conservation of Mass \& Momentum}

\begin{frame}
  \frametitle{Governing Equations: Conservation of Mass \& Momentum}
  \textbf{Conservation of mass}, continuum assumption holds.\\
  Lagrangian (moving with a particle ${}_0\boldsymbol{x}$):
  \begin{equation}
    {^t_0J}\rho(_0\boldsymbol{x},t) = {}\rho(_0\boldsymbol{x},0).
  \end{equation}
Eulerian (outside observer ${}_t\boldsymbol{x}$):
\begin{equation}
  \frac{\partial}{\partial t}\rho(_t\boldsymbol{x},t) = -  \vec{\nabla} \cdot (\rho(_t\boldsymbol{x},t) \boldsymbol{v}(_t\boldsymbol{x},t)).
\end{equation}
Lagrangian and Eulerian view measure differently but give same results. Equations are given in the strong form! \cite{MPM:COURSE}\cite{MIT:CONTINUUM_MECHANICS}
\end{frame}

\begin{frame}
  \textbf{Conservation of momentum}:\\
  Lagrangian (moving with a particle ${}_0\boldsymbol{x}$):
  \begin{equation}
  \rho(_0\boldsymbol{x},0)\boldsymbol{a}(_0\boldsymbol{x},t)
= \vec{\nabla} \cdot \boldsymbol{P}(_0\boldsymbol{x},t) + {\boldsymbol{f}} ^ {\text{body}}(_0\boldsymbol{x},t) {^t_0J}.
  \end{equation}
Eulerian (outside observer ${}_t\boldsymbol{x}$):
\begin{equation}
  \rho(_t\boldsymbol{x},t) \boldsymbol{a}(_t\boldsymbol{x},t) = \vec{\nabla} \cdot \boldsymbol{\sigma}(_t\boldsymbol{x},t) + \boldsymbol{f} ^ {\text{body}}(_t\boldsymbol{x},t)
\end{equation}
Solving this equation will tell us how the velocity fields $\boldsymbol{v}(_t\boldsymbol{x}),\boldsymbol{v}(_0\boldsymbol{x})$ change on the whole domain due to acceleration $\boldsymbol{a}$. This is important to advect particles accounting for all forces. \cite{MPM:COURSE}\cite{MIT:CONTINUUM_MECHANICS}
\end{frame}

\subsection{The Pretty Strong but Mathematically Weak Formulation}
\begin{frame}
\frametitle{The Pretty Strong but Mathematically Weak Formulation}
\begin{align}
  \int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot  \Bigl[(_0\rho_0)(_0\boldsymbol{a}) - & {}_t\boldsymbol{f}^{\text{body}}\Bigr] d _0\boldsymbol{x}= \nonumber \\
										     &  \int _ { \partial \Omega ^ { t^n } } _t\boldsymbol{q} \cdot \boldsymbol{\sigma} d_t\boldsymbol{A} - \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}
\end{align}
\end{frame}
\section{A MPM Guide on GPGPU Pitfalls and Optimizations}

\subsection{c}

\begin{frame}
\frametitle{hi}
  \begin{enumerate}
    \item \cite{MPM:COURSE}
  \end{enumerate}
\end{frame}

\section{Delving Deeper: Further Opportunities}

\subsection{d}

\begin{frame}
\frametitle{hi}
  \begin{enumerate}
    \item \cite{MPM:COURSE}
  \end{enumerate}
\end{frame}

\selectlanguage{english}
\printbibliography
\end{document}
